Proving linearizability using forward simulations\thanks{An extended version of this paper including the missing proofs can be found at \cite{extended}.}

### 2017-05-05 draftCAV/queue.tex

%\begin{figure} [t]
\begin{figure}[t]
\hspace{-10mm}
\begin{minipage}[t]{7.5cm}
\scriptsize{
{	%\centering
	\begin{align*}
%		& \mathsf{Operations}, \mathsf{Values} : \mathsf{Sort} \\
		& \text{\underline{Abstract State}} \\
		& \text{{\ttfamily \bfseries function}}\ \mathsf{present} : \<Ops> \rightarrow \mathbb{B} \\
		& \text{{\ttfamily \bfseries function}}\ \mathsf{pending} : \<Ops> \rightarrow \mathbb{B} \\
		& \text{{\ttfamily \bfseries function}}\ \mathsf{arg} : \<Ops> \rightarrow \<Vals> \\
		& \text{{\ttfamily \bfseries function}}\ \mathsf{before} : \<Ops> \times \<Ops> \rightarrow \mathbb{B} \\
%		& \mathsf{stored}, \mathsf{pending} : \mathsf{Operations} &\mbox{ // the status of an enqueue} \\
%		& \mathsf{arg} : \mathsf{Operations} \rightarrow \mathsf{Values} &\mbox{ // the argument of an enqueue} \\
	\end{align*}
}}
\end{minipage}
\begin{minipage}[t]{7.5cm}
\begin{lstlisting}
void enq($\<Vals>$ x):
  atomic rule inv(enq,x,k = new)
  atomic rule ret(enq,k)

$\<Vals>$ deq():
  atomic rule inv(deq,k = new)
  atomic rule lin(deq,y,k)
  atomic rule ret(deq,y,k)

@@

\begin{minipage}[t]{7.5cm}
\vspace{-3mm}
\begin{lstlisting}
rule inv(enq,x,k):
  $\mathsf{present}$(k) := true
  $\mathsf{pending}$(k) := true
  $\mathsf{arg}$(k) := x
  forall k1 with $\mathsf{present}$(k1)$\land \neg \mathsf{pending}$(k1):
    $\mathsf{before}$(k1,k) := true

rule lin(deq,y,k):
  choose a $\in$ $\mathbb{B}$
  if $($a$\vee\forall k'. \mathsf{present}(k')$=false$) \land \forall k'. \mathsf{present}(k')\Rightarrow \mathsf{pending}(k')$:
    y := EMPTY
  else:
    choose k1 $\in$ $\<Ops>$ with $\mathsf{present}$(k1)$\land\neg \exists k'. \mathsf{present}(k')\land \mathsf{before}(k',$k1$)$
    $\mathsf{present}$(k1) := false
    y := $\mathsf{arg}$(k1)

\end{lstlisting}
\end{minipage}
\begin{minipage}[t]{5cm}
\vspace{-3mm}
\begin{lstlisting}
rule ret(enq,k):
  $\mathsf{pending}$(k) := false

rule inv(deq,k):
  skip

rule ret(deq,y,k):
  return y

>>>>

%\begin{figure} [t]
\begin{figure}[t]
\hspace{-10mm}
\begin{minipage}[t]{7.5cm}
\begin{lstlisting}
void enq($\<Vals>$ v):
  atomic rule inv(enq,v)
  atomic rule lin(enq,v)
  atomic rule ret(enq,v)

\end{lstlisting}
\end{minipage}
\begin{minipage}[t]{7.5cm}
\begin{lstlisting}
$\<Vals>$ deq():
  let v with ...
  atomic rule inv(deq,v)
  atomic rule lin(deq,v)
  atomic rule ret(deq,v)
  return v

@@

\begin{minipage}[t]{7.5cm}
\vspace{-3mm}
\begin{lstlisting}
function present: $\mathbb{O} \to \mathbb{B}$
function pending: $\mathbb{O} \to \mathbb{B}$
function arg: $\mathbb{O} \to \mathbb{V}$
function before: $\mathbb{O} \times \mathbb{O} \to \mathbb{B}$

rule inv(enq,v):
  let k = new $\mathbb{O}$
  present(k) := true
  pending(k) := true
  arg(k) := v
  forall k' with present(k')$\land \lnot$pending(k'):
    before(k',k) := true

rule lin(enq,v):
  pass

rule ret(enq,v):
  let k with arg(k) = v
  pending(k) := false

\end{lstlisting}
\end{minipage}
\begin{minipage}[t]{5cm}
\vspace{-3mm}
\begin{lstlisting}
rule inv(deq,v):
  pass

rule lin(deq,v):
  if v = EMPTY:
    forall k with present(k):
      assert pending(k)
  else:
    assert present(k)
    let k with arg(k) = v
    forall k' with present(k'):
      assert $\lnot$before(k', k)
    present(k) := false

rule ret(deq,v):
  pass
